Indoor Dining Shutdowns Negatively Affect Toronto Restaurants’ Performance and Business Practices

### 2021-02-27 outputs/paper/paper.Rmd
Readjusted figure sizes

This study was conducted for three months, from October 1st to December 31st, 2020 where we studied the comparison between a control group and an intervention group. The intervention group had to partake in the study by their business as a take-out and delivery only restaurant, while the control group remained fully operational with indoor dining services available. This was done intentionally to maintain the integrity of the study by studying the differences between restaurants that closed their indoor dining services and the ones that did not. Survey data was collected via Google Forms to measure the effects of the intervention on restaurant’s employee numbers and revenue, and if changes were made in menu prices and operation days and hours during the intervention period. Responses were processed and analyzed in R [@citeR], and a complete list of packages and their usages can be found in the Appendix. Based on survey responses, it was revealed that most restaurants in the intervention group who shut down their indoor dining services experienced a decrease in their total sales revenue compared to the previous quarter, with many reporting an overall decline of 50% or more. During the study period, restaurants in the intervention group made several changes to their restaurant practices to account for these changes such as reducing the number of employees, adjusting their menu prices, and reducing their hours and days of operation. Meanwhile, the restaurants in the control group who continued operating at full capacity remained relatively stable in terms of their revenue, and did not make any significant changes to their employee counts, menu prices, or operational periods. This suggests that shutdowns of indoor dining practices due to COVID-19 have large implications on restaurant performance, where the revenue, hours and days of operation, and menu prices all get impacted to adjust to deviations from the restaurant’s original business model.
```{r}

@@

```{r samplechar, fig.width=9, fig.height=7, echo = FALSE, message = FALSE, fig.cap="Sample Characteristics of Restaurants by Condition"}
#combining plots for q1-4
grid.arrange(q1, q2, q3, q4)
```

```{r Q10, echo = FALSE, message = FALSE, fig.cap = "Food Courier Service Restaurants Received the Most Orders From During Study Period", fig.width=5, fig.height=3.5}
### Between Oct 1 and Dec 31, which food courier service did you receive the most orders from?
### UberEats
### SkiptheDishes 
### DoorDash
### Other
### In-house delivery service
### None
ggplot(data = all_simulated_dataset, mapping = aes(x = Q11, fill = type)) +
  geom_bar(position = "stack") + coord_flip() +
  scale_x_discrete(limit = c("UberEats", "SkiptheDishes", "DoorDash", "In-house delivery service", "Other", "None")) +
  labs(x = "Delivery Service",
       y = "# of Restaurants",
       title = "Food Courier Service Popularity",
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) + 
  theme_fivethirtyeight() +
  theme(title=element_text(size=9))
```

@@

@@

In terms of effects on revenue for restaurants during the intervention period, Figure \@ref(fig:reveffects) shows us that more restaurants within the treatment group saw decreases in revenue compared to the prior quarterly revenue in which they were offering their full services (July 1 - Sep 30), while the control group's revenue largely remained the same compared to their previous quarterly revenue. In addition to this, for restaurants that saw a decrease in revenue, the intervention group saw the harshest declines (declines of around 50%) in revenue during the intervention period. A Chi-Square test for whether the restaurant saw a decrease in revenue showed that restaurants in the treatment group were more likely to report a decrease in revenue *X*^2^ (2, *N* = 330) = 220.73, *p* < 2.2e-16. The percentage decrease in revenues also showed a significant relationship, where restaurants in the treatment group are most likely to report higher loss of revenue during the intervention period *X*^2^ (3, *N* = 330) = 178.33, *p* < 2.2e-16. Prior to the intervention period, it appears as though both groups had similar breakdowns of revenue which came from take-out and/or delivery as shown in Figure \@ref(fig:Q9).

@@

```{r reveffects, echo = FALSE, message = FALSE, fig.height=5.5, fig.width=5, fig.cap="Changes in Revenue During Intervention Period Compared to the Previous Quarter"}
grid.arrange(q7, q8)
```

@@

@@

```{r Q11, fig.height = 3.5, fig.width = 5, echo = FALSE, message = FALSE, fig.cap = "Adjustments Made to Menu Prices During the Intervention Period"}
### Between (start date of intervention) and (end date of intervention), did you make any adjustments to your menu prices?
### Yes, prices were raised
### Yes, prices were lowered
### No changes were made
ggplot(data = all_simulated_dataset, mapping = aes (x = Q10, fill = type)) +
  geom_bar() +  
  coord_flip() + 
  scale_x_discrete(limit = c("Yes, prices were raised", "Yes, prices were lowered", "No changes were made"), labels = c("Prices were raised", "Prices were lowered", "No changes were made")) + 
  labs(x = "Adjustments to Menu Prices",
       y = "# of Restaurants",
       title = str_wrap("Adjustments to Menu Prices", 30),
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) +
  theme_fivethirtyeight() +
  theme(title=element_text(size=9))

@@

```{r opshourseffects, fig.height=3.5, fig.width=8, echo = FALSE, message = FALSE, fig.cap="Effects on Hours of Operation During Intervention Period by Condition"}
grid.arrange(q12, q13, ncol = 2)
```

```{r opsdayseffects, fig.height=3.5, fig.width=8, echo = FALSE, message = FALSE, fig.cap="Effects on Days of Operation During Intervention Period by Condition"}
grid.arrange(q14, q15, ncol = 2)
```

>>>>

This study was conducted for three months, from October 1st to December 31st, 2020 where we studied the comparison between a control group and an intervention group. The intervention group had to partake in the study by their business as a take-out and delivery only restaurant, while the control group remained fully operational with indoor dining services available. This was done intentionally to maintain the integrity of the study by studying the differences between restaurants that closed their indoor dining services and the ones that did not. Survey data was collected via Google Forms to measure the effects of the intervention on restaurant’s employee numbers and revenue, and if changes were made in menu prices and operation days and hours during the intervention period. Responses were processed and analyzed in R [@citeR], and a complete list of packages and their usages can be found in the Appendix. Based on survey responses, it was revealed that most restaurants in the intervention group who shut down their indoor dining services experienced a decrease in their total sales revenue compared to the previous quarter, with many reporting an overall decline of 50% or more. During the study period, restaurants in the intervention group made several changes to their restaurant practices to account for these changes such as reducing the number of employees, adjusting their menu prices, and reducing their hours and days of operation. Meanwhile, the restaurants in the control group who continued operating at full capacity remained relatively stable in terms of their revenue, and did not make any significant changes to their employee counts, menu prices, or operational periods. This suggests that shutdowns of indoor dining practices due to COVID-19 have large implications on restaurant performance, where the revenue, hours and days of operation, and menu prices all get impacted to adjust to deviations from the restaurant’s original business model.

@@

```{r samplechar, fig.width=8.5, fig.height=7, echo = FALSE, message = FALSE, fig.cap="Sample Characteristics of Restaurants by Condition"}
#combining plots for q1-4
grid.arrange(q1, q2, q3, q4)
```

```{r Q10, echo = FALSE, message = FALSE, fig.cap = "Food Courier Service Restaurants Received the Most Orders From During Study Period", fig.width=5, fig.height=3}
### Between Oct 1 and Dec 31, which food courier service did you receive the most orders from?
### UberEats
### SkiptheDishes 
### DoorDash
### Other
### In-house delivery service
### None
ggplot(data = all_simulated_dataset, mapping = aes(x = Q11, fill = type)) +
  geom_bar(position = "stack") + coord_flip() +
  scale_x_discrete(limit = c("UberEats", "SkiptheDishes", "DoorDash", "In-house delivery service", "Other", "None")) +
  labs(x = "Delivery Service",
       y = "# of Restaurants",
       title = "Food Courier Service Popularity",
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) + 
  theme_fivethirtyeight() +
  theme(title=element_text(size=9))
```

@@

Table \@ref(tab:Q5Q6tTest) displays changes in the numbers of employees working at restaurants in the sample prior to the intervention period (Oct 1) and following the intervention period (Dec 31). This shows us that initially, both the control and treatment group had a very similar number of employees on average. Following the intervention period, however, the treatment group appears to have 10.26 fewer employees on average. Two-tailed Welch's t-tests comparing the two groups' mean employees shows that the difference prior to the intervention period is not a significant finding \newline(*t*(330) = -8.56, *p* > .05), but the difference in employee numbers between the two groups post-intervention is significant(*t*(330) = 19.46, *p* < .05).

@@

Regarding revenue for restaurants during the intervention period, Figure \@ref(fig:reveffects) shows us that more restaurants within the intervention group saw decreases in revenue compared to the prior quarterly revenue in which they were offering their full services (July 1 - Sep 30), while the control group's revenue largely remained the same compared to their previous quarterly revenue. In addition to this, for restaurants that saw a decrease in revenue, the intervention group saw the harshest declines (declines of around 50%) in revenue during the intervention period. A Chi-Square test for whether the restaurant saw a decrease in revenue showed that restaurants in the treatment group were more likely to report a decrease in revenue *X*^2^ (2, *N* = 330) = 220.73, *p* < 2.2e-16. The percentage decrease in revenues also showed a significant relationship, where restaurants in the treatment group are most likely to report higher loss of revenue during the intervention period *X*^2^ (3, *N* = 330) = 178.33, *p* < 2.2e-16. Prior to the intervention period, it appears as though both groups had similar breakdowns of revenue which came from take-out and/or delivery as shown in Figure \@ref(fig:Q9).

@@

```{r reveffects, echo = FALSE, message = FALSE, fig.height=4.7, fig.width=5, fig.cap="Changes in Revenue During Intervention Period Compared to the Previous Quarter"}
grid.arrange(q7, q8)
```

@@

Pearson’s Chi-squared (*X*^2^) tests were conducted to see if there is a statistically significant difference between responses regarding restaurant adjustments in the intervention and control conditions. The test results show that the p-value of each of those adjustments is less than .001, meaning there is a significant relationship between condition (whether or not indoor dining services were shut down or not) and adjustments made to menu prices, hours of operation, and days of operations. That is, restaurants in the intervention condition were more likely to change their menu prices and reduce their operational hours and days.

@@

```{r Q11, fig.height = 3, fig.width = 5, echo = FALSE, message = FALSE, fig.cap = "Adjustments Made to Menu Prices During the Intervention Period"}
### Between (start date of intervention) and (end date of intervention), did you make any adjustments to your menu prices?
### Yes, prices were raised
### Yes, prices were lowered
### No changes were made
ggplot(data = all_simulated_dataset, mapping = aes (x = Q10, fill = type)) +
  geom_bar() +  
  coord_flip() + 
  scale_x_discrete(limit = c("Yes, prices were raised", "Yes, prices were lowered", "No changes were made"), labels = c("Prices were raised", "Prices were lowered", "No changes were made")) + 
  labs(x = "Adjustments to Menu Prices",
       y = "# of Restaurants",
       title = str_wrap("Adjustments to Menu Prices", 30),
       fill = "Condition") +
  scale_fill_tableau(labels = c("Control", "Intervention"), guide = guide_legend(reverse = TRUE)) +
  theme_fivethirtyeight() +
  theme(title=element_text(size=9))

@@

```{r opshourseffects, fig.height=3.3, fig.width=8, echo = FALSE, message = FALSE, fig.cap="Effects on Hours of Operation During Intervention Period by Condition"}
grid.arrange(q12, q13, ncol = 2)
```

```{r opsdayseffects, fig.height=3.3, fig.width=8, echo = FALSE, message = FALSE, fig.cap="Effects on Days of Operation During Intervention Period by Condition"}
grid.arrange(q14, q15, ncol = 2)
```
